Supporting Information : The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-11-26 Paper.Rmd
edited Data section

<!--
Considered datasets ranged widely from population ecology, community ecology , evolutionary and biogeochemical sphere. 
We used four datasets in each category of bivariate (soil C and N data, BMR data, Anolis data, Ceder-creek experiment data) and multivariate (aphid-abundance data, aphid-firstflight data, plankton-abundance data and methane-flux data) examples. Details of each dataset used were given in \ref{SM-Data_detailed}.-->

For bivariate analysis we used four datasets which are briefly described as follows;
\textbf{(i)} to explore the dependence of **soil organic C** and total **soil N** across a wide geographic range (conterminous US), ecosystem type, and traceable methodology we used the RaCA (Rapid Carbon Assessment)
database [@Wills2014, **or Wills and Loecke pending???**]. Specifically, the soil organic C and total soil N
stocks (Mg C or N m$^{-2}$ ) to 1m of soil depth from the central pedon of the 5930 sites were included in this
study. Then, we used extensive set of basal metabolic rate data for \textbf{(ii)} birds [@Mcnab_2009] and \textbf{(iii)} mammals [@Mcnab_2008] as our second and third bivariate datasets to explore the dependence between log(BMR) and log(mass) of respective species. \textbf{(iv)} We used [plant above ground biomass](https://www.cedarcreek.umn.edu/research/data/dataset?ple120) and [plant species percent cover](https://www.cedarcreek.umn.edu/research/data/dataset?pce120) data from a biodiversity experiment [@Tilman2001; @Tilman2006] to calculate plant-aboveground biomass and Shannon's diversity index (H), averaged over each plot (9m by 9m), for year 2000.

Similarly, we applied multivariate analysis on four spatial datasets; \textbf{(i)} aphid-count and \textbf{(ii)} aphid-firstflight date were extracted from the Rothamsted Insect Survey suction-trap data [@Macaulay1988, @Harrington2014]. These two annual timesries (1976-2010) measurements are from 11 locations acroos UK for 20 aphid species (Table \ref{SM-tab_aphid_info}). We used the same data format as used by @Sheppard2016 for their analysis on the impact of climate on aphids' spatial synchrony. In our case, for aphid-abundance data and for aphid-firstflight data, only 10 locations which had atleast 30 years of data (see Figure \ref{SM-fig_aphid_map}) were considered. \textbf{(iii)} Plankton-count data collected at 26 locations over the areas of the North Sea and British seas for 22 species (Table \ref{SM-tab_plankton_info}) and for 56 years (1958-2013). These data were obtained from the Continuous Plankton Recorder (CPR) data set and maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS).
<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).--> [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}). \textbf{(iv)} Finally, methane ($CH_{4}$) fluxes between the soil or water surface and the troposphere were measured at the Great Miami Wetland Mitigation Bank ($36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.

>>>>

\noindent We used datasets that pertain to multiple fields and levels of organization within or related to ecology, including environmental, species-trait, phenological, population, community, and ecosystem functioning data. Copula structure describing the relationship between atmospheric weather variables such as X and Y has been examined previously in the meteorological literature (e.g., REF). We therefore examined environmental variables from the soil instead, using the Rapid Carbon Assessment database (RaCA; @Wills2014).
<!--***DAN: Add Wills and Loecke pending reference if published in time--> 
The database comprises measurements of soil organic carbon and total soil nitrogen (milligrams C or N per m$^{2}$ of soil surface) 
<!--***DAN: Shya, this used to say Mg C, but what is Mg? Do you mean mg for miligram? I have written it out, please check and correct if necessary. Also, I added per m^2 of soil surface, please check.-->
from X 
<!--***DAN: Shya, make this an autolink to the number of data points - 5930, but should be an autolink, so if the number changes it updates-->
locations across the coterminous United States (Fig. X for sampling locations).
<!--***DAN: Shya, please add a sup mat figure showing the sampling locations and a link to it here in place of the X-->
Measurments were from the central pedon of a soil core, to 1 m depth.
<!--***DAN: Terry to check the wording and accuracy here-->
Species-trait data were average species basal metabolic rate (units) and body mass (units) data for X species of birds [@Mcnab_2009] and Y species of mammals [@Mcnab_2008].
<!--***DAN: Shya to add units, and replace X and Y with autolinks that give the numbers-->
These data have been much studied, but to our knowledge the copula structure of the dependence between metabolic rate and body mass has not been examined. 
Species-trait data such as these reflect the coevolution of the two traits.
Phenological data were aphid first-flight dates from 10 locations (Fig. \ref{SM-fig_aphid_map}) 
<!--***DAN: Shya to add a figure shwoing the locations on a map of the UK and cite it here with an automatic citation-->
across the United Kingdom (UK) for 20 aphid species (Table \ref{SM-tab_aphid_info}), for the 35 years from 1976 to 2010. These time series were computed from the the Rothamsted Insect Survey suction-trap dataset [@Macaulay1988, @Harrington2014]. The first of our two population-level datasets was also derived from the RIS suction-trap data, and comprised total counts of aphids trapped for the same locations, species, and years; these are indicators of flying aphid abundance. Data pre-processing for the aphid data was the same as that of @Sheppard2016,
<!--***DAN: Shya, let's discuss please - if we say this we have to be sure we did *exactly* the same pre-processing as Lawrence, whereas I seem to remember we did something slightly different-->
but we exlcuded one of the 11 sites they used so as to include only sites with at least 30 years data.  Our second population dataset comprised average annual abundance estimates for 14 locations (Fig. \ref{SM-fig_plankton_map})
in the North Sea and British seas for 22 plankton taxa (Table \ref{SM-tab_plankton_info}) for the 56 years from 1958 to 2013. The locations are $2^\circ$ by $2^\circ$ grid cells. These data were obtained from the Continuous Plankton Recorder (CPR) dataset maintained by the Sir Alistair Hardy Foundation for Ocean Science (SAHFOS), now part of the Marine Biological Association of the UK. Data pre-processing was similar but not identical to that of @Sheppard2017; see Appendix \ref{SM-Data_details} for details.
Community-level data were plant aboveground biomass and Shannon's diversity index for X plots, each 9m by 9m, maintained at Cedar Creek Ecosystem Science Reserve and sampled in the year 2000 [@Tilman2001; @Tilman2006]. 
Finally, ecosystem functioning data were methane (CH$_{4}$) fluxes between the soil or water surface and the troposphere, measured at 13 locations at daily to weekly intervals from September 2015 to September 2016 at the Great Miami Wetland mitigation bank in Trotwood, Ohio [@Holland1999; @Jarecke2016]. Each included location was measured on at least 50 dates. Sampling locations spanned a hydrologic gradient from upland to seasonal wetland.
See Appendix \ref{SM-Data_details} for additional data details.

The environmental, species-trait, and community datasets, as we studied them, are henceforth called the "bivariate" datasets in the sense that they comprise two quantities measured at different locations or for different species. The phenology and population datasets are "multivariate" in that they comprise, for each taxon, measurements at multiple locations; copula structure can be studied for each pair of locations. Likewise the the ecosystem functioning (methane flux) dataset was  multivariate in this sense.

### 2018-11-26 SupportingInformation.Rmd
edited Data section

>>>>

# Details of the data \label{Data_details}

<!--***DAN: This is a dumping group at the moment, but needs to be organized into a section that gives all the necessary details about the data-->

<!--The CPR is a device which is towed across the ocean behind a ship, filtering plankton from seawater with a continuously
unwinding silk ribbon. The ribbon was cut into 4 inch segments, each corresponding to a particular time and
location in the ocean. Organisms were identified and counted (possibly with sub-sampling) microscopically [@Colebrook1979]. All the samples available for a given variable in a given area in a given month of a given year were
averaged, producing a monthly time series for each year. Where data for a given month was unavailable the missing value was replaced with a median value for that month (drawn from all years in which data was available).-->

 [@Sheppard2017] converted monthly values to yearly values by averaging over all months. But here we only allowed maximum 4 months in a year with missing data which were replaced by median value for that month (drawn from all years in which data was available). We used data from 14 locations which had atleast 45 years time series data (see Figure \ref{SM-fig_plankton_map}).

 The Great Miami Wetland Mitigation Bank  was at $36^{\circ}46^{'}51^{''}$ N, $84^{\circ}20^{'}26^{''}$ W) in Trotwood, Montgomery County, Ohio. Additional information on the history of the site and soil types is detailed in @Jarecke2016. Briefly, the closed static flux chamber method was used [@Holland1999] to measure $CH_{4}$ fluxes at 28 locations across a hydrologic gradient (upland to seasonal wetland) at daily to weekly interval from September 2015 to September 2016. **See and Smyth et al, pending for
additional details???** on greenhouse gas flux methods for this site. We considered 13 locations that had atleast 50 dates data.
